---
title: "Functions"
author: "Iñaki Urrutia"
date: "8/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create a function "impares" that given an array returns the number of odd elements it contains / Crea una función "impares" que dado un vector devuelva la cantidad de elementos impares que contiene.

```{r}
impares <- function(v)
{
  sum(v %% 2 == 1)
}
numbers <- 1:10

impares(numbers)
```

##Create a function "cambio" that, given a matrix of integer numbers, replaces all NA by 0 / Crea una función "cambio" que dada una matriz de numeros enteros reemplaze todos los NA por el valor 0.

```{r}
cambio <- function(m)
{
  m[is.na(m)] = 0
  m
}

m <- matrix(c(1:5, NA, 7:10, NA, 12:16), ncol = 4)
m <- cambio(m)

m
```

##Create a function "unir" that given two arrays returns a new array with the elements of both arrays without repetition / Crea una función "unir" que dados dos vectores devuelva un nuevo vector con los elementos de ambos vectores sin repetidos.

```{r}
unir <- function (v1, v2)
{
  v <- unique(c(v1, v2))
  v
}
v1 <- 1:10
v2 <- 5:15
v <- unir(v1, v2)

v
```

##Create a function "vyc" that given a string returns a list of two components, one containing the vowels and the other the consonants / Crea una función "vyc" que dado un string devuelva una lista de dos componentes que contenga las vocales y las consonantes.
```{r}
vyc <- function (string)
{
  string <- gsub("[^a-zA-Z]", "", string)
  mask_vocales <- as.vector(gregexpr(text = string, pattern = "[aeiou]")[[1]])
  vocales <- strsplit(string, split = "")[[1]][mask_vocales]
  consonantes <- strsplit(string, split = "")[[1]][-mask_vocales]
  list(vocales, consonantes)

}
string = "Esto es una frase de prueba"
res <- vyc(string)

res
```

##Create a function that given an array v and two values x and y (with y an optional parameter), returns a new array with the values included after the first x upto the first y. If y is not indicated, the function will return the values after the first x until the end of the array / Crea una función partir que dado un vector v y dos valores x e y (siendo y opcional), retorne un vector con los valores que aparecen luego del primer x y hasta el primer y. De no indicarse el valor de y se devolveran todos los valores que aparecen luego del primer x hasta el final del vector.

```{r}
subvector <- function(v, x, y = NULL)
{
  ini = which(v == x)[1]
  if (is.null(y))
    fin = length(v)
  else
    fin = which(v == y)[1]
  
  v <- v[ini:fin]
  v
}  
v = 10:1
x = 6
subv <- subvector(v = 10:1, x = 6)

subv
```