---
title: "Input/Output"
author: "Iñaki Urrutia"
date: "8/10/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ask the user to introduce with the kwyboard a string s and a number n and show in the screen n times the string s (without spaces between words) / Pide al usuario que introduzca por teclado un string s y un número n y que muestre en pantalla n veces seguidas el string s (sin espacios entre palabra y palabra). s="hola, n=3, holaholahola

```{r}
s <- readline("Introduce s: ")
n <- readline("Introduce n: ")

print(strrep(s, n))
```

## Create three files called dos.txt, tres.txt and cinco.txt containing the table of 2, 3 and 5 respectively (the first 10 values of each, one number at a time) / Crea tres ficheros llamados dos.txt, tres.txt y cinco.txt que contenga la tabla del 2, la del 3 y la del 5 respectivamente (los primeros 10 valores de cada tabla, un número en cada línea de cada fichero).

## Load the three files created in the last exercise and create a matrix with three columns, one for each file conent / Carga los tres ficheros creados en el punto anterior y construye una matriz que, en cada columna, tengo el contenido de cada fichero.

```{r}
dos = scan("dos.txt", what=numeric())
tres = scan("tres.txt", what=numeric())
cinco = scan("cinco.txt", what=numeric())
a = scan("cinco.txt", what=numeric())

numbers <- matrix(c(dos, tres, cinco), ncol = 3)
numbers
```

## Write the first five rows of the matrix created in the last exercise in a new file called prime.txt and the last five in another file called fin.txt. Both files must have the data separated by commas / Escribe las cinco primera filas de matriz del ejercicio anterior en un fichero nuevo llamado prime.txt y las cinco últimas en otro fichero llamado fin.txt. Ambos ficheros deben tener los datos separados por comas.

```{r}
write.table(numbers[1:5,], "prime.txt", row.names=FALSE, col.names=FALSE, sep = ",")
write.table(numbers[5:nrow(numbers),], "fin.txt", row.names=FALSE, col.names=FALSE, sep = ",")

```

##Given two numbers, f and c, introduced using the keyboard by the user, create a square of f rows and c collumns with character "x". See an example for f=4 and c=3 below: / Dados dos números introducidos por el usuario f y c, crea un cuadrado de f filas y c columnas con el caracter "x". Un ejemplo con f=4 y c=3 sería:
xxx

xxx

xxx

xxx

```{r}
f <- as.numeric(readline("Introduce f: "))
c <- as.numeric(readline("Introduce c: "))
cuadrado <- matrix(data = "x" , nrow = f, ncol = c)

cuadrado
```
