---
title: "Strings"
author: "Iñaki Urrutia"
date: "8/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create a string array contaning your first name and last names (e.g., ["Rocio", "Romero", Zaliz"]). Using that array and the R functions you just learned create a new string with the initial of your first name, a dot, and your last names (e.g., "R. Romero Zaliz") / Crea un vector de strings con tu primer nombre y tus dos apellidos. A partir de éste crea un nuevo string con la inicial de tu nombre (y un punto) y el apellido completo utilizando lo aprendido anteriormente. En mi caso debería quedar: "R. Romero Zaliz".


```{r}
name <- c("Iñaki", "Urrutia", "Sánchez")
paste(paste(substr(name[1],1,1),".", sep = ""), name[2], name[3])
```
## Given an array of strings representing dates (e.g., [“2005-11-28”, “2015-10-18”, “2000-01-01”]), show only those corresponding to odd months / Dado un vector de fechas, expresadas como strings (e.g., [“2005-11-28”, “2015-10-18”, “2000-01-01”]), muestra solamente aquellas correspondientes a los meses impares.


```{r}
dates <- c("2005-11-28", "2015-10-18", "2000-01-01")
months <- as.numeric(substr(dates, 6, 7))
mask <- months%%2 == 1
dates <- dates[mask]

dates

```

## Given a string with several words (e.g., “Esta es una frase, pero no cualquier frase.”) create an array with each of the words in the string (e.g., ["Esta","es","una","frase","pero","no","cualquier","frase"])/ Dado un string con varias palabras (e.g., “Esta es una frase, pero no cualquier frase.”) crea un vector con cada una de las palabras del string (e.g., ["Esta","es","una","frase","pero","no","cualquier","frase"]).

```{r}
words <- "Esta es una frase, pero no cualquier frase."
words <- unlist(strsplit(x = words, split = " "))

words
```

## Search in an array of strings those including only vocals "a" and "e" / Busca las palabras que usan solamente las vocales “a” y “e” en un vector de strings.

```{r}
x <- c("prueba", "elemento", "example", "no")
mask <- unique(append(grep("a", x), grep("e", x)))
mask

x[mask]
```

## Given three numeric arrays representing days, months and years, create an array with dates (only if they are valid) (Hint: research the as.Date function) / Dados tres vectores dia, mes y anno crea un vector con las fechas completas. Si la fecha es inválida, ésta se descartará (Ayuda: investiga la función as.Date).

```{r}
days <- c(23, 20, 05, 32)
months <- c(09, 06, 10, 13)
years <- c(1997, 1980, 2006, 2021)
dates <- paste(months, days, years ,sep = "/")
dates <- dates[!is.na(as.Date(dates, "%m/%d/%y"))]

dates
```